Empirical normative theories

### 2020-08-02 simple.tex
Cosmetic rework of normative analysis

\subsection{Classify propositions}
\begin{axiom}[Normative adequacy]
	\label{ax:norm}
	$\forall i \in I, \phi \in \Phi: 
		[(\exists \ar \in \allargs \suchthat \ar \ileadstost \phi) \land ¬(\exists \ar \in \allargs \suchthat \ar \ileadstost ¬\phi)] ⇒ \phi \in \iPhi \land ¬\phi \notin \iPhi.$
\end{axiom}
This axiom can be simplified, thanks to the following proposition.
\begin{proposition}[Protocol coherence]
	\label{th:protcoh}
	$\forall i \in I, \phi \in \Phi$:
	\begin{equation}
		[\exists \ar \in \allargs \suchthat \ar \ileadstost \phi] ⇒ [\nexists \ar \in \allargs \suchthat \ar \ileadstost ¬\phi].
	\end{equation}
\end{proposition}
\begin{proof}
	Protocol coherence is satisfied by definition of $\ileadstost$. Indeed, if for some $\ar \in \allargs$, $\ar \ileadstost \phi$, then $\forall \ar_1 \in \allargs: (\ar_1, \ar) \nileadsto ¬\phi$; thus, for any $\ar' \in \allargs$, $\ar' \ileadstost ¬\phi$ is false, because it requires that $\forall \ar_1 \in \allargs: (\ar', \ar_1) \ileadsto ¬\phi$, and thus in particular, that $(\ar', \ar) \ileadsto ¬\phi$.
\end{proof}

Therefore, \cref{ax:norm} is actually equivalent to the following property.
\begin{property}[Simple normative adequacy]
	\label{ax:snorm}
	$\forall i \in I, \phi \in \Phi: 
		[\exists \ar \in \allargs \suchthat \ar \ileadstost \phi] ⇒ \phi \in \iPhi \land ¬\phi \notin \iPhi.$
\end{property}

To further analyze this and related matters, given a normative theory and $i \in I$, say that a proposition $\phi \in \Phi$ is \emph{$i$-decidable} iff $\exists \ar \in \allargs \suchthat \ar \ileadstost \phi \lor \ar \ileadstost ¬\phi$.
The $i$-decidable propositions are those on which the normative theory permits to take a position. It will generally be unknown in practical applications, and may be empty.
The next property ensures an unambiguous interpretation of the content of $\iPhi$ for all $i$-decidable propositions. In this definition, $\oplus$ denotes the exclusive disjunction; thus $\phi \in \iPhi \oplus ¬\phi \in \iPhi$ is equivalent to: $\phi \in \iPhi ⇔ ¬\phi \notin \iPhi$. It can be seen to hold by recalling that \cref{ax:snorm} is equivalent to \cref{ax:norm}.

>>>>

\subsection{Classify propositions}
\begin{axiom}[Normative adequacy]
	\label{ax:norm}
	$\forall i \in I, \phi \in \Phi: 
		[(\exists \ar \in \allargs \suchthat \ar \ileadstost \phi) \land ¬(\exists \ar \in \allargs \suchthat \ar \ileadstost ¬\phi)] ⇒ \phi \in \iPhi \land ¬\phi \notin \iPhi.$
\end{axiom}

\Cref{ax:norm} is actually equivalent to the following property.
\begin{property}[Simple normative adequacy]
	\label{ax:snorm}
	$\forall i \in I, \phi \in \Phi: 
		[\exists \ar \in \allargs \suchthat \ar \ileadstost \phi] ⇒ \phi \in \iPhi \land ¬\phi \notin \iPhi.$
\end{property}

To see why this hold, consider the following definition.
\begin{definition}[Protocol coherence]
	\label{def:protcoh}
	$\forall i \in I, \phi \in \Phi$:
	\begin{equation}
		[\exists \ar \in \allargs \suchthat \ar \ileadstost \phi] ⇒ [\nexists \ar \in \allargs \suchthat \ar \ileadstost ¬\phi].
	\end{equation}
\end{definition}
\begin{proposition}
	\label{th:protcoh}
	Any normative theory satisfies Protocol coherence.
\end{proposition}
\begin{proof}
	Protocol coherence is satisfied by definition of $\ileadstost$. Indeed, if for some $\ar \in \allargs$, $\ar \ileadstost \phi$, then $\forall \ar_1 \in \allargs: (\ar_1, \ar) \nileadsto ¬\phi$; thus, for any $\ar' \in \allargs$, $\ar' \ileadstost ¬\phi$ is false, because it requires that $\forall \ar_1 \in \allargs: (\ar', \ar_1) \ileadsto ¬\phi$, and thus in particular, that $(\ar', \ar) \ileadsto ¬\phi$.
\end{proof}

That \cref{ax:norm} is equivalent to \cref{ax:snorm} immediately follows from \cref{th:protcoh}.
\begin{proposition}
	\label{th:normeqsnorm}
	Any normative theory satisfies Normative adequacy iff it satisfies simple normative adequacy.
\end{proposition}

To further analyze this and related matters, given a normative theory and $i \in I$, say that a proposition $\phi \in \Phi$ is \emph{$i$-decidable} iff $\exists \ar \in \allargs \suchthat \ar \ileadstost \phi \lor \ar \ileadstost ¬\phi$.
The $i$-decidable propositions are those on which the normative theory permits to take a position. It will generally be unknown in practical applications, and may be empty.
The next property ensures an unambiguous interpretation of the content of $\iPhi$ for all $i$-decidable propositions. In this definition, $\oplus$ denotes the exclusive disjunction; thus $\phi \in \iPhi \oplus ¬\phi \in \iPhi$ is equivalent to: $\phi \in \iPhi ⇔ ¬\phi \notin \iPhi$. It can be seen to hold by using \cref{th:normeqsnorm}.
